---
title: "Week 2 Assignment - Movie Ratings"
author: "Emin Allen"
date: "`r Sys.Date()`"
output: html_document
---



```{r setup, include=FALSE}
# Load libraries
library(DBI)
library(RMySQL)
library(knitr) 
knitr::opts_chunk$set(echo = TRUE)
```

## Intro
The purpose of this assignment was to manage and prepare data for downstream analysis. 
This was achieved by taking a move ratings data set and storing it in a MySQL database,
then retrieving it into R to check for any missing data. 

## Connect to SQL database and retrieve table into R dataframe
First, we connect to the SQL database and retrieve the table into an R dataframe. 

# Attempt to connect to SQL database
```{r}
connection <- dbConnect(
  RMySQL::MySQL(),
  dbname = "emin.allen15",
  host = "cunydata607sql.mysql.database.azure.com",
  port = 3306,
  user = "emin.allen15",
  password = "a3d09bd8a2b71f9f"
)

# Retrieve data from database
movie_ratings_df <- dbGetQuery(connection, "SELECT * FROM movie_ratings")

# Close connection
dbDisconnect(connection)
```

## Check for missing data
After the table from SQL is loaded into R, we check for missing values

```{r}
missing_data_check <- any(is.na(movie_ratings_df)) 
missing_data_summary <- colSums(is.na(movie_ratings_df))
```

# Print summary of missing data
```{r}
missing_data_check
missing_data_summary
```

Missing data check should result in FALSE, meaning there are no missing values found anywhere in the dataframe. 

Furthermore, there should be zero missing values in any column. 

## Display table

```{r}
knitr::kable(movie_ratings_df)
```

## Conclusion
No missing data was encountered.

No imputation or removal was necessary. 