---
title: "Week 2 Assignment - Movie Ratings"
author: "Emin Allen"
date: "`r Sys.Date()`"
output: html_document
---



```{r setup, include=FALSE}
# Load libraries
library(DBI)
library(RMySQL)
library(knitr) 
knitr::opts_chunk$set(echo = TRUE)
```

# Intro
The purpose of this assignment was to manage and prepare data for downstream analysis. 
This was achieved by taking a move ratings data set and storing it in a MySQL database,
then retrieving it into R to check for any missing data. 

## Connect to SQL database and retrieve table into R dataframe
First, we connect to the SQL database and retrieve the table into an R dataframe. 

## Attempt to connect to SQL database
```{r}
connection <- dbConnect(
  RMySQL::MySQL(),
  dbname = "emin.allen15",
  host = "cunydata607sql.mysql.database.azure.com",
  port = 3306,
  user = "emin.allen15",
  password = "a3d09bd8a2b71f9f"
)

# Retrieve data from database
movie_ratings_df <- dbGetQuery(connection, "SELECT * FROM movie_ratings")

# Close connection
dbDisconnect(connection)
```

## Check for missing data
After the table from SQL is loaded into R, we check for missing values

```{r}
missing_data_check <- any(is.na(movie_ratings_df)) 
missing_data_summary <- colSums(is.na(movie_ratings_df))
```

## Print summary of missing data
```{r}
missing_data_check
missing_data_summary
```

Missing data check should result in FALSE, meaning there are no missing values found anywhere in the dataframe. 

Furthermore, there should be zero missing values in any column. 

# Missing values can be handled with simple removal or imputation.

## Simple removal of rows from the dataframe that contain missing values
```{r}
  movie_ratings_clean <- na.omit(movie_ratings_df)
```

## Imputation by mean, median, mode, or constant value
```{r}
movie_ratings_clean <- movie_ratings_df
```

### Mean imputation
```{r}
movie_ratings_clean$rating[is.na(movie_ratings_clean$rating)] <-
mean(movie_ratings_clean$rating, na.rm = TRUE)
```

### Median imputation
```{r}
movie_ratings_clean$rating[is.na(movie_ratings_clean$rating)] <-       median(movie_ratings_clean$rating, na.rm = TRUE)
```

### Mode imputation
```{r}
mode_rating <- names(sort(table(movie_ratings_clean$rating), decreasing = TRUE))[1]

movie_ratings_clean$rating[is.na(movie_ratings_clean$rating)] <- as.numeric(mode_rating)
```
  
  
### Constant value imputation
```{r}
movie_ratings_clean$person[is.na(movie_ratings_clean$person)] <- "Unknown"
```


### Display table
```{r}
knitr::kable(movie_ratings_df)
```

## Conclusion
No missing data was encountered.

No imputation or removal was necessary. 